---
title: "MAF_analysis"
author: "Jinyu Wang"
date: "April 23, 2017"
output: html_document
---
###THe object of this part is to generate Fig.3A and Fig.4A of the paper

Load the ggplot2 library 

```{r}
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(grid)

```

Set the working directory and load into the data

```{r}
setwd("C:/Users/jinyuw/Desktop/BCB546X-Project/PEB_Power_Final_Project/Jinyu/MAF_data")
MAF <- read.delim("SNP_MAF.txt",header = TRUE)
head(MAF)
summary(MAF)
```
Change the name of each group so that it matches with the label of the plot on the paper 
```{r}
MAF$Group <- as.character(MAF$Group)
MAF$Group[MAF$Group == "Processing"] <- "Proc"
MAF$Group[MAF$Group == "Fresh_market"] <- "FM"
MAF$Group[MAF$Group == "Vintage"] <- "Vint"
MAF$Group <- as.factor(MAF$Group)
```


Reorder the level of group to make it to be the same order as what it shown on the paper 
```{r}
MAF$Group <- factor(MAF$Group, levels =c("Proc", "FM", "Vint", "Pimp"))
```

theme(axis.text.y =element_text(size=12))+
Subset the data for chromosome 1 and Plot the figure for chromosome 1.

```{r}
MAF_Chr1 <- MAF[MAF$Chromosome == 1, ]
dim(MAF_Chr1)
p1 <- ggplot(MAF_Chr1, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR1", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 80))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p1

```



```{r}
MAF_Chr2 <- MAF[MAF$Chromosome == 2, ]
dim(MAF_Chr2)
p2<- ggplot(MAF_Chr2, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR2", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 80))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p2
```


```{r}
MAF_Chr3 <- MAF[MAF$Chromosome == 3, ]
dim(MAF_Chr3)
p3<- ggplot(MAF_Chr3, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR3", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 80))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p3
```


```{r}
MAF_Chr4 <- MAF[MAF$Chromosome == 4, ]
dim(MAF_Chr4)
p4<- ggplot(MAF_Chr4, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR4", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 80))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p4
```

```{r}
MAF_Chr5 <- MAF[MAF$Chromosome == 5, ]
dim(MAF_Chr5)
p5 <- ggplot(MAF_Chr5, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR5", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 80))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p5
```

```{r}
MAF_Chr6 <- MAF[MAF$Chromosome == 6, ]
dim(MAF_Chr6)
p6 <- ggplot(MAF_Chr6, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR6", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 80))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p6
```



```{r}
MAF_Chr7 <- MAF[MAF$Chromosome == 7, ]
dim(MAF_Chr7)
p7 <- ggplot(MAF_Chr7, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR7", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 80))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p7 
```
```{r}
MAF_Chr8 <- MAF[MAF$Chromosome == 8, ]
dim(MAF_Chr8)
p8 <- ggplot(MAF_Chr8, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR8", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 80))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p8
```


```{r}
MAF_Chr9 <- MAF[MAF$Chromosome == 9, ]
dim(MAF_Chr9)
p9 <- ggplot(MAF_Chr9, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR9", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 80))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p9
```

```{r}
MAF_Chr10 <- MAF[MAF$Chromosome == 10, ]
dim(MAF_Chr10)
p10 <- ggplot(MAF_Chr10, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR10", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 100))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p10
```



```{r}
MAF_Chr11 <- MAF[MAF$Chromosome == 11, ]
dim(MAF_Chr11)
p11 <- ggplot(MAF_Chr11, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR11", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 100))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p11
```

```{r}
MAF_Chr12 <- MAF[MAF$Chromosome == 12, ]
dim(MAF_Chr12)
p12 <- ggplot(MAF_Chr12, aes(x=Position, y=MAF, fill = MAF, colour="red")) +
  geom_bar(stat="identity", position = "dodge", show.legend=F) + 
  facet_wrap(~Group, scales = "free_y", nrow = 4, ncol=1, strip.position = "left") + 
  theme_minimal()+
  theme(plot.title = element_text( face="bold"))+
  labs(title = "CHR12", x = "Physical map distance (Mbp)", y = "MAF")+
  scale_x_continuous(breaks=c(0,20,40, 60, 100))+
  scale_y_continuous(breaks=c(0.20, 0.60, 1.00))+
  coord_cartesian(ylim = c(0,1), xlim = c(0, 91))
p12
```


Using the grid.arrange() to print out the figure for prensetation. MAF of every three chromosomes are ploted on the same page.

```{r}

pdf("CHR1-2-3.pdf")
grid.arrange(  p1, p2, p3, ncol = 1, heights = c(.3,.3, 0.3), widths = .4)
dev.off()
```

```{r}

pdf("CHR4-5-6.pdf")
grid.arrange(  p4, p5, p6, ncol = 1, heights = c(.3,.3, 0.3), widths = .4)
dev.off()
```



```{r}

pdf("CHR7-8-9.pdf")
grid.arrange(  p7, p8, p9, ncol = 1, heights = c(.3,.3, 0.3), widths = .4)
dev.off()
```



```{r}

pdf("CHR10-11-12.pdf")
grid.arrange(  p10, p11, p12, ncol = 1, heights = c(.3,.3, 0.3), widths = .4)
dev.off()
```

