---
title: "PCA_Analysis"
output:
  html_document: default
  pdf_document: default
---
### PCA Analysis
1. Load necessary libraries
2. Import the imputed data file created from imputation_workflow
3. Make sure col and row names are set properly, and remove if necessary
```{r Loading data}
library(tidyr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(grid)
library(lattice)
setwd("C:/Users/mdzievit.IASTATE/Dropbox/Classes/EEOB_546X/PEB_Power_Final_Project/Structure_Analysis/")
data <- read.delim("imputed_genetic_map_data_EXPIM2012_wh.txt",header = TRUE)
rownames(data) <- data[,1]
data <- data[,-1]
```
4. Run PCA analysis. Paper did not give a lot of details regarding their methods and settings they used.
  - Set 2 new variables x and VAR. These are the PC scores (x) and the variance explained by each of the PCs
```{r PCA Analysis}
pca.data <- prcomp(data,center = TRUE)
x.pca.data <- pca.data$x
x.pca.data <- as.data.frame(x.pca.data)
var.pca.data <- summary(pca.data)
var.pca.data <- var.pca.data$importance
var.pca.data <- var.pca.data[2,]
```

## Import genotype information to make PCA plots
# Plot #1 using all sub-populations
1. Import the genotype information file. Skipping the first line as it is just a descriptor that it is a table. Checked the head information in gitbash
  - Also need to rename and add the genotype name column in the x.pca.data so we can join them later
```{r  Extracting Data}
genotype <- read.delim("Table_S1.txt",header = TRUE, skip = 1)
colnames(genotype)
genotype <- genotype %>% select(SolCAP.T.number,Market.Class)
x.pca.data$SolCAP.T.number <- rownames(x.pca.data)


##Identified incorrect classifications in the genotype file related to the original PCA figure. Need to correct these to make the right plot. Fresh market/Processing is one that was changed after. Vintage/processing were ones that were changed after analysis. I am changing these to their correct classification to remove the type. It causes problems down stream.
levels(genotype$Market.Class)
genotype[genotype == "Fresh market/Processing"] <- "Processing"
genotype[genotype == "Vintage/Processing"] <- "Processing"
genotype <- droplevels(genotype)
levels(genotype$Market.Class)
```
2. Join the genotype information file with the PC file
```{r Joining genotype file}
info.x.pca.data <- genotype %>% 
  right_join(x.pca.data, by = c("SolCAP.T.number")) %>%
  droplevels()

##Reducing the number of PCs to 3, keeping name and market class and export that for future analysis
info.x.pca.data.red <- info.x.pca.data %>% select(SolCAP.T.number,Market.Class,PC1,PC2,PC3)
write.table(info.x.pca.data.red,file = "full_PC.txt",col.names = TRUE,sep = "\t", quote = FALSE)
rm(info.x.pca.data.red)
```

3. Plot the data to look like the figure from the paper
```{r Plotting the first PCA figure}
##Determines the levels of the market
levels(info.x.pca.data$Market.Class)

##Identified incorrect classifications in the genotype file related to the original PCA figure. Need to correct these to make the right plot. Fresh market/Processing is one that was changed after so it will be color coded as fresh market. Vintage/processing were ones that were changed after analysis, so will be color coded as processing

##Colors and shapes match the paper
colors <- c("violet","Blue","Gold","Red","Green","Black","Gray")
shapes <- c(0,1,4,2,3,5,22)

figure1.a <- ggplot(aes(x = PC1, y = PC2),data = info.x.pca.data) + 
  geom_point(aes(colour = Market.Class, shape = Market.Class),size = 2) +
  scale_color_manual(values = colors) +
  scale_shape_manual(values = shapes) + 
  labs(x = paste("PC1 (",var.pca.data[1]*100,"%)",sep = ""),
       y = paste("PC2 (", var.pca.data[2]*100,"%)",sep = ""),
       title = "Figure 1. A",
       subtitle = "PCA based on 4,393 SNP markers: All Sub-populations") +
  theme_bw() +
  theme(aspect.ratio = 1)
```

# Plot#2 Using data for only 3 large-fruited cultivated sub-populations
- This consists of processing, fresh market, and vintage
1. Need to subset the data for the different types for the analysis
```{r Subsetting the data for second PCA}
data.info <- data
data.info$SolCAP.T.number <- rownames(data.info) 
data.info <- genotype %>% 
  right_join(data.info, by = c("SolCAP.T.number")) %>%
  droplevels()

data.filt <- data.info %>%
  filter(Market.Class == "Fresh market" | Market.Class == "Processing" | Market.Class == "Processing" | Market.Class == "Vintage") %>%
  select(-Market.Class) %>%
  droplevels()

rownames(data.filt) <- data.filt$SolCAP.T.number
data.filt$SolCAP.T.number <- NULL

```
2. Run PCA on the new dataset and subset important information
```{r Running 2nd PCA Plot }
pca2.data.filt <- prcomp(data.filt,center = TRUE)
x.pca2.data.filt <- pca2.data.filt$x
x.pca2.data.filt <- as.data.frame(x.pca2.data.filt)
var.pca2.data.filt <- summary(pca2.data.filt)
var.pca2.data.filt <- var.pca2.data.filt$importance
var.pca2.data.filt <- var.pca2.data.filt[2,]
```

3. Recombine with the genotype information
```{r Joining with the Genotyping file}
x.pca2.data.filt$SolCAP.T.number <- rownames(x.pca2.data.filt)
info.x.pca2.data.filt <- genotype %>% 
  right_join(x.pca2.data.filt, by = c("SolCAP.T.number")) %>%
  droplevels()

##Reducing PCA2 dataset, and export it for future analysis
info.x.pca2.data.filt.red <- info.x.pca2.data.filt %>% select(SolCAP.T.number,Market.Class,PC1,PC2,PC3)
write.table(info.x.pca2.data.filt.red,file = "Proc_PC.txt",col.names = TRUE,sep = "\t", quote = FALSE)
rm(info.x.pca2.data.filt.red)
```

4. Plot the figure using the details from the paper
```{r Plot of the 2nd PCA}
levels(info.x.pca2.data.filt$Market.Class)
colors2 <- c("Blue","Red","Green")
shapes2 <- c(1,2,3)
figure1.b <- ggplot(aes(x = PC1, y = PC2),data = info.x.pca2.data.filt) + 
  geom_point(aes(colour = Market.Class, shape = Market.Class),size = 2) +
  scale_color_manual(values = colors2) +
  scale_shape_manual(values = shapes2) + 
  labs(x = paste("PC1 (",var.pca2.data.filt[1]*100,"%)",sep = ""),
       y = paste("PC2 (", var.pca2.data.filt[2]*100,"%)",sep = ""),
       title = "Figure 1. B",
       subtitle = "PCA based on 4,393 SNP markers: 3 large fruited sub-populations") +
  theme_bw() +
  theme(aspect.ratio = 1)

```
5. Create a joint figure for output
```{r}
##Run this outside of markdown
pdf("figure1.pdf")
figure1.a
figure1.b
dev.off()
pdf("figure1_both_horz.pdf",onefile = TRUE, paper = 'A4r', width = 11, height = 8) 
grid.arrange(figure1.a, figure1.b,ncol = 2,heights = (c(.5,.5)))
dev.off()
```
