---
title: 'Structure Figures: Full Data'
output:
  html_document: default
  html_notebook: default
---
###This file will input the .Q files from admixture and create new figures from the data. Additionally, it will import data to recreate the PCA plots for comparison.

### Need to load the necessary libraries
```{r Loading necessary libraries}
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
library(grid)
library(lattice)
```

### Importing the admixture results from k = 5 and k = 7
```{r Data import}
###Import the full dataset PC information and import the sample_ID file with market class
setwd("C:/Users/mdzievit.IASTATE/Dropbox/Classes/EEOB_546X/PEB_Power_Final_Project/Structure_Analysis/Admixture_Results/")
subpop5 <- read.table("tomato.5.Q")
subpop7 <- read.table("tomato.7.Q")
sampID_full <- read.table("tomato_sampID_info.txt", sep = "\t",header = TRUE)


```

### Renaming columns and finding the maximum ancestry
```{r Rename and max ancestry}
##Rename the columns
colnames(subpop5)
colnames(subpop5) <- c("Pop1","Pop2","Pop3","Pop4","Pop5")
colnames(subpop5)

##ID the max value and then which subpop it comes from
subpop5$max <- apply(subpop5,1,max)
subpop5$Pop <- colnames(subpop5)[apply(subpop5,1,which.max)]

##Do the same for the k = 7
colnames(subpop7)
colnames(subpop7) <- c("Pop1","Pop2","Pop3","Pop4","Pop5","Pop6","Pop7")
colnames(subpop7)

subpop7$max <- apply(subpop7,1,max)
subpop7$Pop <- colnames(subpop7)[apply(subpop7,1,which.max)]
```

### Importing the PCs from the structure_workflow documentation. See that for how that file was created. Combining this information with out .q files and identifying the admixture lines
```{r Import PCs and combine files}
##Need to import these PCA information
full_PC <- read.table("full_PC.txt", sep = "\t")

##Combine files
subpop5.info <- cbind(sampID_full,subpop5)
subpop7.info <- cbind(sampID_full,subpop7)
rm(subpop5,subpop7)

##Determining if there is admixture and classifying it accordingly
subpop5.info <- subpop5.info %>% mutate(NewPop = ifelse(max >= .51,Pop,"Admix"))
subpop7.info <- subpop7.info %>% mutate(NewPop = ifelse(max >= .51,Pop,"Admix"))
```

### Plotting the figures. Kept the colors and shapes the same from the original figure. The original figure colors and shapes are the same, however, with k = 5 and k = 7, I tried to match as best as possible.
```{r k equal 5 PC plot }
##Creating 5 subpop PC plot to compare with original
full_PC.subpop5 <- subpop5.info %>%
  select(SolCAP.T.number,NewPop) %>%
  left_join(full_PC,by = "SolCAP.T.number")

colors <- c("violet","Blue","Gold","Red","Green","Black","Gray")
shapes <- c(0,1,4,2,3,5,22)

colors2 <- c("Orange","Blue","Black","Green","Gray","Red")
shapes2 <- c(13,1,5,3,0,2)

origPC_Plot <- ggplot(aes(x = PC1, y = PC2),data = full_PC.subpop5) + 
  geom_point(aes(colour = Market.Class, shape = Market.Class),size = 3) +
  scale_color_manual(values = colors) +
  scale_shape_manual(values = shapes) + 
  theme_bw() +
  theme(aspect.ratio = 1,legend.position = "bottom",plot.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
  labs(title = "Original figure, subpopulations assigned by breeders")

structure5_PC_Plot <- ggplot(aes(x = PC1, y = PC2),data = full_PC.subpop5) + 
  geom_point(aes(colour = NewPop, shape = NewPop),size = 3) +
  scale_color_manual(values = colors2) +
  scale_shape_manual(values = shapes2) + 
  theme_bw() +
  theme(aspect.ratio = 1,legend.position = "bottom",plot.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
  labs(title = "Subpopulation colored according to Admixture (K=5)")

##Output these
#pdf("PC_Comp-Subpop5.pdf",onefile = TRUE, paper = 'A4r', width = 11, height = 8)
grid.arrange(origPC_Plot,structure5_PC_Plot,ncol = 2)
#dev.off()
```

###K = 7 plot and compare with original plot
```{r k equal 7 PC plot }
full_PC.subpop7 <- subpop7.info %>%
  select(SolCAP.T.number,NewPop) %>%
  left_join(full_PC,by = "SolCAP.T.number")

colors <- c("violet","Blue","Gold","Red","Green","Black","Gray")
shapes <- c(0,1,4,2,3,5,22)

colors3 <- c("Orange","Red","Gray","Black","Red","Blue","Green","Red")
shapes3 <- c(13,17,0,5,2,1,3,25)

structure7_PC_Plot <- ggplot(aes(x = PC1, y = PC2),data = full_PC.subpop7) + 
  geom_point(aes(colour = NewPop, shape = NewPop),size = 3) +
  scale_color_manual(values = colors3) +
  scale_shape_manual(values = shapes3) + 
  theme_bw() +
  theme(aspect.ratio = 1,legend.position = "bottom",plot.title = element_text(size = 10),
        legend.text = element_text(size = 12)) +
  labs(title = "Subpopulation colored according to Admixture (K=7)")

##Output files
#pdf("PC_Comp-Subpop7.pdf",onefile = TRUE, paper = 'A4r', width = 11, height = 8)
grid.arrange(origPC_Plot,structure7_PC_Plot,ncol = 2)
#dev.off()
```


### Comparing the k = 5 and k = 7 PC plots
```{r Comparing k equals 5 and 7 plots }
##Output files
#pdf("PC_Comp_both.pdf",onefile = TRUE, paper = 'A4r', width = 11, height = 8)
grid.arrange(structure5_PC_Plot,structure7_PC_Plot,ncol = 2)
#dev.off()
```

### Going to make barplots that summarize the ancestry for each genotype
```{r Making bar plots for k equals 5}

#####Make the structure barplot figure for subpop5
subpop5.tidy <- subpop5.info %>%
  arrange(NewPop,desc(Pop1),desc(Pop2),desc(Pop3),desc(Pop4),desc(Pop5)) %>%
  select(-Market.Class,-max,-Pop) %>%
  mutate(ind = c(1:dim(subpop5.info)[1])) %>%
  gather(SubPop,Ancestry,-ind,-SolCAP.T.number,-NewPop)

bar_color <- c("Blue","Black","Green","Gray","Red")
five_subpop <- ggplot(aes(x = ind,y = Ancestry,fill = SubPop), data = subpop5.tidy) + 
  geom_bar(stat = "identity",width = 1) +
  scale_fill_manual(values = bar_color) +
  facet_wrap(~NewPop, strip.position = "left", scales = "free_y", ncol = 1) +
  theme(panel.spacing = unit(0, "lines"), 
        strip.background = element_blank(),
        strip.placement = "outside") + 
  theme_classic() +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), 
        strip.text.y = element_text(angle = 180), panel.spacing = unit(0,"lines")) + 
  coord_flip() +
  labs(subtitle = "Admixture k = 5")
```

### We are combings the subpop 5 and subpop7 files. We are doing this because we want to preserve the order of the individuals from the subpop5 so we can see how the subpopulations change from k = 5 to k = 7
```{r Making barplots for k equals 7}
#####Make the structure barplot figure for subpop7
subpop7.tidy.noind <- subpop7.info %>%
  arrange(NewPop,desc(Pop1),desc(Pop2),desc(Pop3),desc(Pop4),desc(Pop5),
          desc(Pop6),desc(Pop7)) %>%
  select(-Market.Class,-max,-Pop) %>%
  gather(SubPop,Ancestry,-NewPop,-SolCAP.T.number)

##Combine the original order of the subpop5 dataframe with the subpop7. We want to keep the original order of the inds so we can see how they change when we increase k = 7
subpop7.ordered5 <- subpop5.tidy %>%
  select(SolCAP.T.number,ind,NewPop) %>%
  left_join(subpop7.tidy.noind,by = c("SolCAP.T.number")) %>%
  unique()


##The reds are labelled differently to differentiate it. This is because these sub-pops are all from the same original subpopulation in the original dataset.
bar_color2 <- c("Red","Gray","Black","RosyBrown","Blue","Green","FireBrick4")

seven_subpop2 <- ggplot(aes(x = ind,y = Ancestry,fill = SubPop), data = subpop7.ordered5) + 
  geom_bar(stat = "identity",width = 1) +
  facet_wrap(~NewPop.x,scales = "free_y", ncol = 1) +
  scale_fill_manual(values = bar_color2) +
  theme_classic() +
  theme(panel.spacing = unit(0, "lines"), 
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  coord_flip() +
  labs(subtitle = "Admixture k = 7")

##Output these files
#pdf("Comp_5_7_admixture.pdf",onefile = TRUE, paper = 'A4r', width = 11, height = 8)
grid.arrange(five_subpop, seven_subpop2,ncol = 2,
             top = "Comparison of k = 5 to k = 7 classification")
#dev.off()
```



